\section{Conclusion}
\label{sec:conclusion}

We presented a multi-agent simulation framework in which LLM-driven provider and insurer agents negotiate prior authorization for a specialty medication under conditions of information asymmetry, where each agent has access to different clinical and policy documents. In a 3$\times$3 factorial experiment varying provider and insurer strategies on a single clinically ambiguous case, we found that the insurer's binary choice at initial review (request additional documentation or deny outright) is the mechanism through which gray-zone ambiguity is resolved.

Default insurer agents, those given no explicit strategic instructions, denied the drug while approving diagnostic tests, applying coverage policy literally rather than interpretively. This behavior suggests that AI-based adjudication systems deployed without explicit interpretive guidance may default to denial when clinical evidence is ambiguous. When both agents lacked strategic guidance (the NP\_NI condition), the outcome was complete denial achieved through literal policy application by the insurer and exhaustive but futile appeals by the provider, reproducing the worst cooperative outcome without adversarial intent.

Game-theoretic analysis shows that the strategic structure of the interaction is parameter-dependent. At the narrow, directly observed administrative costs (\$5.79 per turn for providers versus \$0.05 per turn for insurers, a 116-fold asymmetry), the interaction constitutes an asymmetric exploitation game in which denial is nearly costless for the insurer. However, these observed costs omit factors that constrain real-world insurer behavior (litigation risk, reputational damage, network-exit costs, and patient-outcome externalities) whose true magnitudes are not directly measurable in our simulation. Under richer cost models that incorporate these factors, the sensitivity analysis shows that the interaction transitions to a prisoner's dilemma in which mutual cooperation is Pareto-dominant (both sides would be better off cooperating), yet both still face individual incentives to defect. The analysis identifies the cost threshold at which this transition occurs, providing a quantitative target for policy interventions aimed at raising the effective cost of denial.

Future work should extend this framework to multiple clinical cases, coverage policies, foundation models, and replicated runs to assess the generalizability of these findings and to identify conditions under which cooperative equilibria can be sustained.
