\section{Discussion}
\label{sec:discussion}

\subsection{Strategic Interpretation in the Gray Zone}
\label{sec:gray-zone}

The simulation results show that when clinical facts and coverage policy are held constant, the patient's treatment outcome is determined entirely by how agents interpret ambiguity. The pivotal decision is whether the insurer treats ambiguity as an invitation for dialogue (\texttt{pending\_info}) or as grounds for denial. Cooperating insurers chose dialogue and ultimately approved the drug; defecting and default insurers chose denial and never engaged. This single binary choice---request information or deny---resolved the gray zone across all nine conditions, suggesting that the mechanism of coordination failure in AI-mediated prior authorization is not adversarial strategy per se, but the absence of interpretive flexibility. Default insurer agents, given no strategic guidance, applied coverage policy literally and produced denial in every case (Section~\ref{sec:default-insurer}), implying that AI adjudication systems deployed without explicit interpretive instructions will default to denial whenever clinical evidence does not unambiguously satisfy policy criteria.

\subsection{Game Structure and the Prisoner's Dilemma Question}
\label{sec:game-structure}

We now return to the motivating question posed by \citet{moritz2025lancet}: does mutual AI adoption in prior authorization produce a prisoner's dilemma? The answer depends on parameters whose real-world values are not directly observed in our simulation.

The payoff analysis (Section~\ref{sec:payoff}) shows that the game's classification hinges on the true cost of insurer denial. Our simulation captures only the narrow electronic transaction cost (\$0.05 per insurer turn, from CAQH), which produces a 116-fold asymmetry with the provider's per-turn cost (\$5.79). At these values, denial is nearly free for the insurer, and the game is an asymmetric exploitation structure rather than a prisoner's dilemma: the insurer has a dominant strategy to defect, and the provider's strategy choice is irrelevant once the insurer does so. From the provider's perspective, this is worse than a prisoner's dilemma---there is no coordination failure to resolve because one side simply dominates.

However, the \$0.05 figure captures only direct processing costs and omits factors that constrain real-world insurer behavior: litigation exposure from denied claims, reputational damage from public denial-rate reporting, provider network attrition from systematic denials, and patient-outcome externalities from delayed treatment. The sensitivity analysis (Section~\ref{sec:payoff}) shows that when the insurer's effective per-turn costs rise to incorporate these factors---approaching the provider's cost level---the game crosses into prisoner's dilemma territory at moderate values of the burden-weight parameter $\alpha_I$. In this regime, mutual cooperation becomes Pareto-dominant (both sides would be better off cooperating), yet both still face individual incentives to defect, reproducing the coordination failure that \citet{moritz2025lancet} hypothesized.

Whether real-world insurer costs already reach this threshold is an empirical question that our simulation cannot resolve. But the default-agent results point toward its practical relevance. The NP\_NI condition---where neither agent received strategic guidance---produced the same deadlock as mutual defection (zero services approved), not through adversarial intent but through literal policy application by the insurer and exhaustive procedural appeals by the provider (Section~\ref{sec:default-insurer}). If unguided AI systems already reproduce the worst cooperative outcome through bureaucratic attrition, the question is not whether a coordination failure exists, but whether its structure is one that policy intervention can address. The sensitivity analysis provides a quantitative target: raising the effective cost of insurer denial, through regulatory mechanisms such as mandatory denial-rate reporting, litigation exposure, or accountability for patient outcomes, could shift the game from exploitation to a prisoner's dilemma where mutual cooperation is the Pareto-dominant equilibrium.

% \subsection{Limitations and Future Work}
% \label{sec:limitations}
%
% \paragraph{Utility function incompleteness.}
% Equations~\ref{eq:u_provider}--\ref{eq:u_insurer} capture direct financial outcomes and administrative costs but omit factors that constrain real-world insurer defection: litigation risk, network effects (providers exiting insurer networks after systematic denials), reputational costs (public denial-rate reporting affecting enrollment), and patient outcome externalities (denied treatments producing downstream emergency utilization). These omitted factors all punish insurer defection, explaining why our model finds defection lucrative (\$2,274 at mutual defection) while real-world insurers face meaningful constraints. Incorporating these costs would raise the effective price of insurer defection, potentially crossing the threshold where the game transitions from exploitation to a prisoner's dilemma (Section~\ref{sec:payoff}).
%
% \paragraph{Appeal chain truncation.}
% Our simulation terminates at Level~2 (Independent Review Entity), but real-world providers can pursue further legal appeals through ALJ hearings, the Medicare Appeals Council, and federal court. Truncating the appeal chain removes the provider's credible escalation threat, biasing the game toward insurer dominance. This is distinct from the litigation risk discussed above as an omitted cost; here, the provider's strategic option set is also truncated, not just the insurer's cost function.
%
% \paragraph{Cross-turn service attribution.}
% When the provider resubmits with a changed service set, the audit log records what changed but not \emph{why}. A dropped line could represent a concession, error correction, or strategic gaming; these are observationally equivalent. Our final-state pricing sidesteps this by pricing only terminal services, at the cost of losing signal about within-run concessions. A fix: require the provider to emit a structured \texttt{resubmission\_reason} per changed line, logged but not exposed to the insurer to preserve information asymmetry.
%
% \paragraph{Reviewer differentiation.}
% Real-world review levels involve different reviewer types with different authority, and AI involvement in adjudication likely decreases at higher levels. Our insurer agent adjudicates with the same prompt at all levels (Appendix~\ref{sec:appeal-levels}). Insurers' internal decision processes are largely proprietary, making it difficult to model how criteria shift across levels.
%
% \paragraph{Temporal costs.}
% The simulation measures burden in turns but does not model the cost of time. Each turn represents days to weeks of delay; for a patient with active Crohn's and bowel stricturing, delay risks disease progression and hospitalization. A time-aware utility function would penalize turns as patient harm, likely strengthening the case against prolonged denial strategies.
%
% \paragraph{Single case, single model, single run.}
% This study uses one clinical case with one insurer policy. The gray zone mechanism (cytopenias as steroid contraindication) is case-specific; other drug-disease-policy combinations will produce different ambiguity structures. All agents use GPT-5 via Azure OpenAI; different models may exhibit different strategic sensitivities and error rates (e.g., the Q5105 code hallucination in DP\_CI). Each condition was run once, but LLM outputs are stochastic: alternate runs of CP\_DI shifted the Nash equilibrium from (D,D) to (C,D) (Section~\ref{sec:stochastic}). Within each run, default agent behavior was internally consistent (the default insurer denied at every decision point and never used \texttt{pending\_info}), but this within-run consistency does not preclude across-run variation. Multiple replications across cases, policy types, and foundation models are needed to assess generalizability.
