\section{Introduction}
\label{sec:introduction}
Prior authorization (PA) is a utilization management process in which healthcare providers must obtain approval from an insurer before delivering certain medical services or prescribing specific medications. Insurers adjudicate these requests by approving, modifying, or denying them, or by requesting additional documentation. Providers who receive unfavorable decisions may appeal through successive levels of review. This creates substantial administrative burden: the average physician's practice completes an estimated 39 PA requests per week, spending 13 hours on PA-related tasks \citep{ama2024pa}. Both sides are now deploying AI to gain strategic advantage. Insurers use AI to scrutinize claims and flag requests for denial, while providers deploy AI to draft appeals and optimize clinical documentation \citep{williams2024hfma}.

\citet{moritz2025lancet} argue that this dynamic resembles a prisoner's dilemma, a game-theoretic scenario in which two parties, each acting in rational self-interest, produce a collectively worse outcome than if both had cooperated. They warn that ``each party's rational decision to adopt AI for self-protection creates collectively harmful outcomes for everyone.'' The resulting incentive structure is straightforward: if one party deploys AI, the other must follow suit or face a competitive disadvantage. Yet mutual AI deployment may produce worse outcomes than mutual restraint, including more administrative burden, more denials, and more appeals, without improving patient care. This raises a fundamental question: what happens when both provider and insurer deploy large language models (LLMs) to interpret ambiguous coverage policies? Does the predicted coordination failure actually emerge, and if so, through what mechanism?

We address these questions through a controlled simulation study. Our contributions are threefold:
\begin{enumerate}
    \item \textbf{Multi-agent PA simulation framework.}
    We develop a simulation environment in which LLM-driven provider and insurer agents negotiate prior authorization through iterative cycles of submission, adjudication, and appeal. The agents operate under information asymmetry (the provider observes only clinical treatment guidelines while the insurer observes only the coverage policy) and their behavioral strategies (cooperative, adversarial, or no explicit guidance) are controlled through prompt-based instructions.
    \item \textbf{3$\times$3 factorial experiment.}
    Using a single clinically ambiguous case (infliximab for Crohn's disease), we cross three provider strategies with three insurer strategies and identify a key decision point: whether the insurer requests additional clinical information (\texttt{pending\_info}) or issues an outright denial. This decision is the primary mechanism through which ambiguity is resolved. Default (unguided) insurer agents apply coverage policy literally, denying ambiguous treatments rather than engaging in cooperative interpretation of clinical evidence.
    \item \textbf{Game-theoretic payoff analysis.}
    We derive utility functions (numerical measures of each agent's outcome quality) from simulation results and show that the game's classification depends on the assumed cost parameters: at observed administrative costs, the 116-fold cost asymmetry between providers and insurers makes denial nearly costless for insurers, producing a one-sided exploitation structure; under richer cost models that account for litigation, reputational, and network-adequacy costs, the game can transition to the prisoner's dilemma hypothesized by prior work.
\end{enumerate}
